---
title: "Figure 5"
site: workflowr::wflow_site
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: console
---

## Model

Let $y_{ij}$ denote the number of individuals with mutation $i$ for cancer type $j$ and $n_{j}$ the total number of individuals (excluding hypermutators). Then

\begin{align*}
y_{ij} &\sim \text{Binomial}(\theta_{ij}, n_{j})\\
\text{logit}(\theta_{ij}) &= \alpha + \beta x\\
\alpha &\sim \text{normal}(0, 10)\\
\beta &\sim \text{normal}(0, 3).\\
\end{align*}
$x$ is an indicator for cancer type $j$ =2.

## Analysis


```{r packages, message=FALSE}
library(hallberg2025.base)
library(gridExtra)
library(grid)
library(tidyverse)
library(dplyr)
library(magrittr)
library(here)
```

This analysis is limited to genes where the overall number of mutations in each of the two comparisons (uterine endometrioid vs ovarian endometrioid and GI mucinous vs ovarianmucinous) is at least 10%.

```{r analytic.data}
ue.v.oe <- "Uterine endometrioid - Ovarian endometrioid" %>%
    str_wrap(25)
gi.v.om <- "GI mucinous - Ovarian mucinous" %>%
    str_wrap(25)
bayes.out <- readRDS(here("output",
                      "07-figure5.rmd",
                      "bayes.rds")) %>%
    filter(gene != "hypermutator") %>%
    unite("uid", c(gene, tumors), remove=FALSE) %>%
    mutate(tumors=str_replace_all(tumors, "vs", "-"),
           tumors=case_when(tumors=="Ovarian endometrioid - Uterine endometrioid"~ue.v.oe,
                            tumors=="Ovarian mucinous - GI mucinous"~gi.v.om,
                            TRUE~tumors),
           tumors=str_wrap(tumors, 25)) %>%
    filter(tumors %in% c(ue.v.oe, gi.v.om))
bayes <- filter(bayes.out, parameter=="beta")
stats2 <- bayes %>%
    mutate(lor=mean,
           low=`5%`,
           high=`95%`)
```

# Log odds ratio figures

```{r endo.mafs}
lr.data <- filter(stats2, tumors==ue.v.oe)
base.size <- 16
```

```{r lrfig}
bfig <- lr.data %>%
    ggplot(aes(lor, gene)) +
    geom_errorbarh(aes(xmin=low, xmax=high,
                       color=pathway),
                   height=0.25,
                   linewidth=1.2) +
    geom_point(aes(color=pathway), size=5) +
    theme_classic(base_size=base.size) +
    theme(panel.grid=element_blank(),
          strip.background=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          legend.direction="vertical",
          legend.text=element_text(size=24)) +
    geom_vline(xintercept=0, linetype="dashed",
               color="gray") +
    ylab("") + xlab("Log odds ratio") +
    coord_cartesian(xlim=c(-10, 10)) +
    guides(color=guide_legend(title="", ncol=2,
                              override.aes=list(size=4)))
lor.leg <- ggpubr::get_legend(bfig)
bfig <- bfig + guides(color="none")
```

```{r lor.grobs}
endo.lor <- ggplotGrob(bfig)
lr.data <- filter(stats2, tumors==gi.v.om)
<<lrfig>>
muc.lor <- ggplotGrob(bfig)
```


```{r nsignif}
e <- 0.1
nsignif <- stats2 %>%
    mutate(is.signif=ifelse(low + e > 0 | high - e < 0, 1, 0)) %>%
    group_by(tumors) %>%
    summarize(n=sum(is.signif))
```

# Mutation frequency figures

```{r colors}
dx.colors <- tumor_colors()
endo.colors <- dx.colors[c("Uterine endometrioid",
                           "Ovarian endometrioid")]
muc.colors <- dx.colors[c("Ovarian mucinous",
                          "Colorectal mucinous")]
names(muc.colors)[2] <- "GI mucinous"
```

```{r prevalence}
bayes.out2 <- bayes.out %>%
    mutate(tmp=str_replace_all(parameter, "\\[", ""),
           tmp=str_replace_all(tmp, "\\]", ""),
           parameter=tmp) %>%
    select(-tmp) %>%
    filter(tumors %in% c(ue.v.oe, gi.v.om))
cancer.ref <- filter(bayes.out2, parameter=="theta1") %>%
    select(gene, pathway, ct,
           "5%", "50%", "95%") %>%
    mutate(cancertype=map_chr(ct, function(x) rownames(x)[1]),
           empirical=map_dbl(ct, function(x) x[1,1]/x[1,2]),
           label=map_chr(ct, function(x) paste0(x[1, 1], "/", x[1,2])))
cancer.ref %>%
    filter(gene=="NF1") %>%
    select(gene, label, cancertype)
cancer <- filter(bayes.out2, parameter=="theta2") %>%
    select(gene, pathway, ct,
           "5%", "50%", "95%") %>%
    mutate(cancertype=map_chr(ct, function(x) rownames(x)[2]),
           empirical=map_dbl(ct, function(x) x[2, 1]/x[2, 2]),
           label=map_chr(ct, function(x) paste0(x[2, 1], "/", x[2,2])))
cancer %>%
    filter(gene=="NF1") %>%
    select(gene, label, cancertype)
prev <- bind_rows(cancer.ref, cancer)
prev %>%
    filter(gene=="NF1") %>%
    select(gene, label, cancertype)
dwidth <- 0.6
text_size <- 6
```

```{r endodata}
err <- 0.1
prev2 <- prev %>%
    mutate(comparison=ifelse(grepl("endo", cancertype),
                             "Endometrioid",
                             "Mucinous"))
endo <- filter(prev2, comparison=="Endometrioid") %>%
    mutate(genei=factor(gene),
           breaks=as.numeric(genei),
           e=ifelse(cancertype %in% c("Ovarian endometrioid",
                                      "Ovarian mucinous"),
                    err, -err),
           genei=breaks+e)
dat <- endo
tumor.colors <- endo.colors
```

```{r figcode}
a <- dat %>%
    ggplot(aes(`50%`, genei, label=label)) +
    geom_errorbarh(aes(xmin=`5%`, xmax=`95%`,
                       color=cancertype), height=0.2,
                   linewidth=1) +
##    geom_point(shape="X", aes(empirical, genei), color="gray",
##               size=3) +
    geom_point(aes(color=cancertype),
               size=3) +
    theme_classic(base_size=base.size) +
    theme(panel.grid=element_blank(),
          strip.background=element_rect(fill="white",
                                        color="white"),
          ##axis.text.x=element_text(size=18),
          ##axis.title.x=element_text(size=20),
          legend.direction="vertical",
          legend.text=element_text(size=24)) +
    ylab("") +
    xlab("Mutation frequency") +
    xlim(c(0, 1)) +
    scale_y_continuous(breaks=dat$breaks, labels=dat$gene) +
    scale_color_manual(values=tumor.colors)
b <- a + geom_text(aes(x=`95%`+0.02), data=dat,
                   size=text_size,
                   hjust=0) +
    guides(color=guide_legend(title=""),
           fill=guide_legend(title=""))
```

```{r endofig}
endo.maf.leg <- ggpubr::get_legend(b)
b <- b+guides(color="none", fill="none")
endo.maf <- ggplotGrob(b)
```

```{r mucdata}
muc <- filter(prev2, comparison=="Mucinous") %>%
    mutate(genei=factor(gene),
           breaks=as.numeric(genei),
           e=ifelse(cancertype == "Ovarian mucinous",
                    err, -err),
           genei=breaks+e)
dat <- muc
tumor.colors <- muc.colors
```

```{r mucfig}
<<figcode>>
muc.maf.leg <- ggpubr::get_legend(b)
b <- b+guides(color="none", fill="none")
muc.maf <- ggplotGrob(b)
##print(b)
```

```{r fig5, fig.width=20, fig.height=12, dev=c("pdf", "png")}
lo <- rbind(c(1, 1, 2, 2, 5, 3, 3, 4, 4),
            c(6, 6, 6, 7, 7, 7, 8, 8, 8))
widths <- c(0.5, 0.5, 0.375, .375, 0.15, 0.5, 0.5, 0.375, 0.375)
widths <- unit(widths/sum(widths), "npc")
heights <- c(1, 0.2)
ng <- nullGrob()
grob <- arrangeGrob(endo.maf, endo.lor,
                    muc.maf, muc.lor, ng,
                    endo.maf.leg,
                    lor.leg,
                    muc.maf.leg,
                    #ng,
                    widths=widths,
                    heights=heights,
                    layout_matrix=lo)
vp.overall <- viewport(0.025, 0.025,
                       width=unit(0.95, "npc"), height=unit(0.95, "npc"),
                       just=c("left", "bottom"))
gl <- grid.layout(2, 5, widths=widths, heights=heights)
##grid.show.layout(gl)
grid.newpage()
pushViewport(vp.overall)
grid.draw(grob)
popViewport()
grid.text(c("a", "b"),
          x=unit(c(0.02, 0.52), "npc"),
          y=unit(0.98, "npc"),
          gp=gpar(cex=3))
grid.text("Ovarian and uterine endometrioid",
          x=unit(0.25, "npc"),
          y=unit(0.98, "npc"),
          gp=gpar(cex=1.75))
grid.text("GI mucinous and ovarian mucinous",
          x=unit(0.75, "npc"),
          y=unit(0.98, "npc"),
          gp=gpar(cex=1.75))
```



